---
title: "Hands-on Exercise 4"
description: |
  In this hands-on exercise, you will gain hands-on experience on using appropriate functions of spatstat package to perform spatial point patterns analysis
author:
  - name: Niharika Avula 
    url: https://example.com/norajones
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
      self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Data

The following three data sets will be used:

* CHILDCARE, a point feature data providing both location and attribute information of childcare centres. It was downloaded from Data.gov.sg and is in geojson format.

* MP14_SUBZONE_WEB_PL, a polygon feature data providing information of URA 2014 Master Plan Planning Subzone boundary data. It is in ESRI shapefile format. This data set was also downloaded from Data.gov.sg.

* CostalOutline, a polygon feature data showing the national boundary of Singapore. It is provided by SLA and is in ESRI shapefile format.

# R Packages

* The following five R Packages are used

```{r echo = TRUE, eval = TRUE}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
```

# Spatial Data Wrangling

## Importing the spatial data

```{r echo = TRUE, eval = TRUE}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```
```{r echo = TRUE, eval = TRUE}
sg_sf <- st_read(dsn = "data", layer="CostalOutline")
```

```{r echo = TRUE, eval = TRUE}
mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")
```

## Reference System Information of the geospatial data

Checking for Childcare_sf Dataframe, it is projected in SVY21 with EPSG code 3414

```{r echo = TRUE, eval = TRUE}
st_crs(childcare_sf)
```

Checking for sg_sf Dataframe

```{r echo = TRUE, eval = TRUE}
st_crs(sg_sf)
```
* sg_sf dataframe is projected in SVY21 but the EPSG code is 9001 instead of 3414, hence need to transform to the correct code

```{r echo = TRUE, eval = TRUE}
sg_sf3414 <- st_transform(sg_sf, 3414)
```

```{r echo = TRUE, eval = TRUE}
st_crs(sg_sf3414)
```
Checking for mpsz_sf Dataframe

```{r echo = TRUE, eval = TRUE}
st_crs(mpsz_sf)
```
* mpsz_sf dataframe is projected in SVY21 but the EPSG code is 9001 instead of 3414, hence need to transform to the correct code

```{r echo = TRUE, eval = TRUE}
mpsz_sf3414 <- st_transform(mpsz_sf, 3414)
```

```{r echo = TRUE, eval = TRUE}
st_crs(mpsz_sf3414)
```

## Mapping the geospatial data sets

```{r echo = TRUE, eval = TRUE}
tmap_mode("plot")
qtm(mpsz_sf3414)
```
```{r echo = TRUE, eval = TRUE}
tmap_mode("plot")
qtm(sg_sf3414)
```
```{r echo = TRUE, eval = TRUE}
tmap_mode("plot")
qtm(childcare_sf)
```

```{r echo = TRUE, eval = TRUE}
tmap_mode("plot")
qtm(mpsz_sf3414) + qtm(childcare_sf)
```
This shows that the referencing system and coordinate values of the two data frames are referred to similar spatial context. 

## Preparing Pin Map

```{r echo = TRUE, eval = TRUE}
tmap_mode("view")
qtm(childcare_sf)+
  tm_dots()
```
* Leaflet for R API is being used in the interactive mode
* The advantage of this interactive pin map is it allows us to navigate and zoom around the map freely
* The background of the internet map layer can also be changed

```{r echo = TRUE, eval = TRUE}
tmap_mode('plot')
```
Reminder: Always remember to switch back to plot mode after the interactive map. This is because, each interactive mode will consume a connection. You should also avoid displaying ecessive numbers of interactive maps (i.e. not more than 10) in one RMarkdown document when publish on Netlify.

# Geospatial Data wrangling

## Converting sf data frames to sp’s Spatial* class

This is because there are many geospatial analysis packages that require the input geospatial data in sp’s Spatial* classes. The code chunk below uses as_Spatial() of sf package to convert the three geospatial data from simple feature data frame to sp’s Spatial* class.

```{r echo = TRUE, eval = TRUE}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

```{r echo = TRUE, eval = TRUE}
childcare 
```
```{r echo = TRUE, eval = TRUE}
mpsz
```

```{r echo = TRUE, eval = TRUE}
sg
```

## Converting the Spatial* class into generic sp format

Spatstat requires the analytical data in ppp object form. There is no direct way to convert a Spatial* classes into ppp object. We need to convert the Spatial classes* into Spatial object first.The codes chunk below converts the Spatial* classes into generic sp objects.

```{r echo = TRUE, eval = TRUE}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

```{r echo = TRUE, eval = TRUE}
childcare_sp
```

```{r echo = TRUE, eval = TRUE}
sg_sp
```

## Converting the generic sp format into spatstat’s ppp format

Use as.ppp() function of spatstat to convert the spatial data into spatstat’s ppp object format

```{r echo = TRUE, eval = TRUE}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```

```{r echo = TRUE, eval = TRUE}
plot(childcare_ppp)
```

Summary statistics of the newly created ppp object can be viewed by using the code chunk below
```{r echo = TRUE, eval = TRUE}
summary(childcare_ppp)
```

## Handling duplicated points

There is a warning message about duplicates, which is a significant issue in spatial point pattern analysis

* Checking for duplication in a ppp object

```{r echo = TRUE, eval = TRUE}
any(duplicated(childcare_ppp))
```
To count the number of co-indicence point, we will use the multiplicity() function as shown in the code chunk below.

```{r echo = TRUE, eval = TRUE}
multiplicity(childcare_ppp)
```
If we want to know how many locations have more than one point event, we can use the code chunk below.

```{r echo = TRUE, eval = TRUE}
sum(multiplicity(childcare_ppp) > 1)
```
There are 128 duplicated point events and to view the locations of these duplicate point events, we will plot childcare data by using the code chunk below.

```{r echo = TRUE, eval = TRUE}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```

```{r echo = TRUE, eval = TRUE}
tmap_mode('plot')
```

```{r echo = TRUE, eval = TRUE}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

To spot the duplicate points from the map shown above, jittering technique can be used which will add a small perturbation to the duplicate points so that they do not occupy the exact same space. The code chunk below implements the jittering approach.

```{r echo = TRUE, eval = TRUE}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

```{r echo = TRUE, eval = TRUE}
any(duplicated(childcare_ppp_jit))
```


## Creating owin object

When analysing spatial point patterns, it is a good practice to confine the analysis with a geographical area like Singapore boundary. In spatstat, an object called owin is specially designed to represent this polygonal region.

The code chunk below is used to covert sg SpatialPolygon object into owin object of spatstat.

```{r echo = TRUE, eval = TRUE}
sg_owin <- as(sg_sp, "owin")
```

```{r echo = TRUE, eval = TRUE}
plot(sg_owin)
```
```{r echo = TRUE, eval = TRUE}
summary(sg_owin)
```

## Combining point events object and owin object

In this last step of geospatial data wrangling, we will extract childcare events that are located within Singapore by using the code chunk below.

Should we be using childcare_PPP_jit?

```{r echo = TRUE, eval = TRUE}
childcareSG_ppp = childcare_ppp_jit[sg_owin]
```

The output object combined both the point and polygon feature in one ppp object class as shown below.

```{r echo = TRUE, eval = TRUE}
summary(childcareSG_ppp)
```

```{r echo = TRUE, eval = TRUE}
plot(childcareSG_ppp)
```

# First-order Spatial Point Patterns Analysis

In this section, I will Learn how to perform first-order SPPA by using spatstat package. The hands-on exercise will focus on:

* deriving kernel density estimation (KDE) layer for visualising and exploring the intensity of point processes,
* performing Confirmatory Spatial Point Patterns Analysis by using Nearest Neighbour statisticS

## Kernel Density Estimation

In this section, I will learn how to compute the kernel density estimation (KDE) of childcare services in Singapore.

### Computing kernel density estimation using automatic bandwidth selection method

```{r echo = TRUE, eval = TRUE}
kde_childcareSG_bw <- density(childcareSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian")
```
```{r echo = TRUE, eval = TRUE}
plot(kde_childcareSG_bw)
```
The density values of the output range from 0 to 0.000035 which is way too small to comprehend. This is because the default unit of measurement of svy21 is in meter. As a result, the density values computed is in “number of points per square meter”.

We can retrieve the bandwidth used to compute the kde layer by using the code chunk below

```{r echo = TRUE, eval = TRUE}
bw <- bw.diggle(childcareSG_ppp)
bw
```

### Rescalling KDE values

rescale() is used to covert the unit of measurement from meter to kilometer.

```{r echo = TRUE, eval = TRUE}
childcareSG_ppp.km <- rescale(childcareSG_ppp, 1000, "km")
```

Re-run density() using the rescale data set and plot the output kde map
```{r echo = TRUE, eval = TRUE}
kde_childcareSG.bw <- density(childcareSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG.bw)
```

### Working with different automatic badwidth methods

Beside bw.diggle(), there are three other spatstat functions can be used to determine the bandwidth, they are: bw.CvL(), bw.scott(), and bw.ppl().

```{r echo = TRUE, eval = TRUE}
bw.CvL(childcareSG_ppp.km)
```

```{r echo = TRUE, eval = TRUE}
bw.scott(childcareSG_ppp.km)
```

```{r echo = TRUE, eval = TRUE}
bw.ppl(childcareSG_ppp.km)
```

```{r echo = TRUE, eval = TRUE}
bw.diggle(childcareSG_ppp.km)
```
The code chunk beow will be used to compare the output of using bw.diggle and bw.ppl methods.

```{r echo = TRUE, eval = TRUE}
kde_childcareSG.ppl <- density(childcareSG_ppp.km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```
### Working with different kernel methods

By default, the kernel method used in density.ppp() is gaussian. But there are three other options, namely: Epanechnikov, Quartic and Dics.

The code chunk below will be used to compute three more kernel density estimations by using these three kernel function.

```{r echo = TRUE, eval = TRUE}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```

### Fixed and Adaptive KDE

COMPUTING KDE BY USING FIXED BANDWIDTH of 600m

```{r echo = TRUE, eval = TRUE}
kde_childcareSG_600 <- density(childcareSG_ppp.km, sigma=0.6, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG_600)
```

COMPUTING KDE BY USING ADAPTIVE BANDWIDTH

* Fixed bandwidth method is very sensitive to highly skew distribution of spatial point patterns over geographical units for example urban versus rural. One way to overcome this problem is by using adaptive bandwidth instead.
* We can derive adaptive kernel density estimation by using density.adaptive() of spatstat.

```{r echo = TRUE, eval = TRUE}
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
plot(kde_childcareSG_adaptive)
```
We can compare the fixed and adaptive kernel density estimation outputs by using the code chunk below

```{r echo = TRUE, eval = TRUE}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")
```
### Converting KDE output into grid object

```{r echo = TRUE, eval = TRUE}
gridded_kde_childcareSG_bw <- as.SpatialGridDataFrame.im(kde_childcareSG.bw)
spplot(gridded_kde_childcareSG_bw)
```
### CONVERTING GRIDDED OUTPUT INTO RASTER

```{r echo = TRUE, eval = TRUE}
kde_childcareSG_bw_raster <- raster(gridded_kde_childcareSG_bw)
```

Properties of kde_childcareSG_bw_raster RasterLayer, Notice that the crs property is NA.
```{r echo = TRUE, eval = TRUE}
kde_childcareSG_bw_raster
```

ASSIGNING PROJECTION SYSTEMS

```{r echo = TRUE, eval = TRUE}
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster
```

Displaying the raster in cartographic quality map using tmap package
```{r echo = TRUE, eval = TRUE}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```

## Comparing Spatial Point Patterns using KDE

I will learn how to compare KDE of childcare at Ponggol, Tampines, Chua Chu Kang and Jurong West planning areas.

EXTRACTING STUDY AREA
```{r echo = TRUE, eval = TRUE}
pg = mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
tm = mpsz[mpsz@data$PLN_AREA_N == "TAMPINES",]
ck = mpsz[mpsz@data$PLN_AREA_N == "CHOA CHU KANG",]
jw = mpsz[mpsz@data$PLN_AREA_N == "JURONG WEST",]
```

Plotting target planning areas
```{r echo = TRUE, eval = TRUE}
par(mfrow=c(2,2))
plot(pg, main = "Ponggol")
plot(tm, main = "Tampines")
plot(ck, main = "Choa Chu Kang")
plot(jw, main = "Jurong West")
```
CONVERTING THE SPATIAL POINT DATA FRAME INTO GENERIC SP FORMAT

```{r echo = TRUE, eval = TRUE}
pg_sp = as(pg, "SpatialPolygons")
tm_sp = as(tm, "SpatialPolygons")
ck_sp = as(ck, "SpatialPolygons")
jw_sp = as(jw, "SpatialPolygons")
```

CREATING OWIN OBJECT

```{r echo = TRUE, eval = TRUE}
pg_owin = as(pg_sp, "owin")
tm_owin = as(tm_sp, "owin")
ck_owin = as(ck_sp, "owin")
jw_owin = as(jw_sp, "owin")
```

COMBINING CHILDCARE POINTS AND THE STUDY AREA

```{r echo = TRUE, eval = TRUE}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
```

rescale() function is used to trasnform the unit of measurement from metre to kilometre

```{r echo = TRUE, eval = TRUE}
childcare_pg_ppp.km = rescale(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale(childcare_jw_ppp, 1000, "km")
```

The code chunk below is used to plot these four study areas and the locations of the childcare centres.

```{r echo = TRUE, eval = TRUE}
par(mfrow=c(2,2))
plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```
COMPUTING KDE

* The code chunk below will be used to compute the KDE of these four planning area. bw.diggle method is used to derive the bandwidth of each

```{r echo = TRUE, eval = TRUE}
par(mfrow=c(2,2))
plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Jurong West")
```
COMPUTING FIXED BANDWIDTH KDE

```{r echo = TRUE, eval = TRUE}
par(mfrow=c(2,2))
plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chua Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Jurong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
```
# Nearest Neighbour Analysis

In this section, I will perform the Clark-Evans test of aggregation for a spatial point pattern by using clarkevans.test() of statspat.

The test hypotheses are:

Ho = The distribution of childcare services are randomly distributed.

H1= The distribution of childcare services are not randomly distributed.

The 95% confident interval will be used.

Testing spatial point patterns using Clark and Evans Test

```{r echo = TRUE, eval = TRUE}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```
## Clark and Evans Test: Choa Chu Kang planning area

In the code chunk below, clarkevans.test() of spatstat is used to performs Clark-Evans test of aggregation for childcare centre in Choa Chu Kang planning area.

```{r echo = TRUE, eval = TRUE}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```
## Clark and Evans Test: Choa Chu Kang planning area

In the code chunk below, clarkevans.test() of spatstat is used to performs Clark-Evans test of aggregation for childcare centre in Tampines planning area.

```{r echo = TRUE, eval = TRUE}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

# Second-order Spatial Point Patterns Analysis

## Analysing Spatial Point Process Using G-Function

The G function measures the distribution of the distances from an arbitrary event to its nearest event. In this section, I will learn how to compute G-function estimation by using Gest() of spatstat package. I will also learn how to perform monta carlo simulation test using envelope() of spatstat package

### Choa Chu Kang planning area

COMPUTING G-FUNCTION ESTIMATION

The code chunk below is used to compute G-function using Gest() of spatat package.

```{r echo = TRUE, eval = TRUE}
G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))
```

PERFORMING COMPLETE SPATIAL RANDOMNESS TEST

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with G-fucntion
```{r echo = TRUE, eval = TRUE}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)
```
```{r echo = TRUE, eval = TRUE}
plot(G_CK.csr)
```

### Tampines planning area

COMPUTING G-FUNCTION ESTIMATION

The code chunk below is used to compute G-function using Gest() of spatat package.

```{r echo = TRUE, eval = TRUE}
G_tm = Gest(childcare_tm_ppp, correction = "best")
plot(G_tm)
```

PERFORMING COMPLETE SPATIAL RANDOMNESS TEST

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected is p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r echo = TRUE, eval = TRUE}
G_tm.csr <- envelope(childcare_tm_ppp, Gest, correction = "all", nsim = 999)
```
```{r echo = TRUE, eval = TRUE}
plot(G_tm.csr)
```
## Analysing Spatial Point Process Using F-Function

The F function estimates the empty space function F(r) or its hazard rate h(r) from a point pattern in a window of arbitrary shape. In this section, I will learn how to compute F-function estimation by using Fest() of spatstat package. I will also learn how to perform monta carlo simulation test using envelope() of spatstat package.

### Choa Chu Kang planning area

```{r echo = TRUE, eval = TRUE}
F_CK = Fest(childcare_ck_ppp)
plot(F_CK)
```

Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with F-fucntion


```{r echo = TRUE, eval = TRUE}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 999)
```

```{r echo = TRUE, eval = TRUE}
plot(F_CK.csr)
```

### Tampines planning area

```{r echo = TRUE, eval = TRUE}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```
PERFORMING COMPLETE SPATIAL RANDOMNESS TEST

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected is p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r echo = TRUE, eval = TRUE}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 999)
```

```{r echo = TRUE, eval = TRUE}
plot(F_tm.csr)
```

## Analysing Spatial Point Process Using K-Function

K-function measures the number of events found up to a given distance of any particular event. In this section, I will learn how to compute K-function estimates by using Kest() of spatstat package. I will also learn how to perform monta carlo simulation test using envelope() of spatstat package.

### Choa Chu Kang planning area

```{r echo = TRUE, eval = TRUE}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```
PERFORMING COMPLETE SPATIAL RANDOMNESS TEST

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r echo = TRUE, eval = TRUE}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```
```{r echo = TRUE, eval = TRUE}
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

### Tampines planning area

COMPUTING K-FUCNTION ESTIMATION

```{r echo = TRUE, eval = TRUE}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

PERFORMING COMPLETE SPATIAL RANDOMNESS TEST

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r echo = TRUE, eval = TRUE}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```
```{r echo = TRUE, eval = TRUE}
plot(K_tm.csr, . - r ~ r, 
     xlab="d", ylab="K(d)-r", xlim=c(0,500))
```


## Analysing Spatial Point Process Using L-Function

In this section, I will learn how to compute L-function estimation by using Lest() of spatstat package. I will also learn how to perform monta carlo simulation test using envelope() of spatstat package.

### Choa Chu Kang planning area

COMPUTING L FUCNTION ESTIMATION

```{r echo = TRUE, eval = TRUE}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```
PERFORMING COMPLETE SPATIAL RANDOMNESS TEST

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value if smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r echo = TRUE, eval = TRUE}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r echo = TRUE, eval = TRUE}
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
### Tampines planning area

COMPUTING L-FUCNTION ESTIMATE

```{r echo = TRUE, eval = TRUE}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```
PERFORMING COMPLETE SPATIAL RANDOMNESS TEST

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below will be used to perform the hypothesis testing.

```{r echo = TRUE, eval = TRUE}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r echo = TRUE, eval = TRUE}
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
```
