---
title: "Hands-on Exercise 2"
description: |
 In this hands-on exercise, I will learn how to handle geospatial data in R by using sf package.
author:
  - name: Niharika Avula 
    url: https://example.com/norajones
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r echo = TRUE, eval = TRUE}
packages <- c('sf', 'tidyverse')
for (p in packages){
  if (!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
  }
```

# Importing Geospatial Data

## Importing polygon feature data in shapefile format

The code chunk below uses st_read() function of sf package to import MP14_SUBZONE_WEB_PL shapefile into R as a polygon feature data frame. 

```{r echo = TRUE, eval = TRUE}
mpsz = st_read(dsn =  "Data/Geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

## Importing polyline feature data in shapefile form

The code chunk below uses st_read() function of sf package to import CyclingPath shapefile into R as line feature data frame.

```{r echo = TRUE, eval = TRUE}
cyclingpath = st_read(dsn = "data/geospatial", 
                         layer = "CyclingPath")
```

## Importing GIS data in kml format

The code chunk below will be used to import the kml into R. Notice that in the code chunk below, the complete path and the kml file extension were provided

```{r echo = TRUE, eval = TRUE}
preschool = st_read("data/geospatial/pre-schools-location-kml.kml")
```

# Checking the Content of a Simple Feature Data Frame

## Working with st_geometry()
```{r echo = TRUE, eval = TRUE}
st_geometry(mpsz)
```

## Working with glimpse()

glimpse() report reveals the data type of each fields.

```{r echo = TRUE, eval = TRUE}
glimpse(mpsz)
```

## Working with head()

```{r echo = TRUE, eval = TRUE}
head(mpsz, n=5) 
```

# Plotting the Geospatial Data

```{r echo = TRUE, eval = TRUE}
plot(mpsz)
```

```{r echo = TRUE, eval = TRUE}
plot(st_geometry(mpsz))
```

```{r echo = TRUE, eval = TRUE}
plot(mpsz["PLN_AREA_N"])
```

# Working with Projection

```{r echo = TRUE, eval = TRUE}
st_crs(mpsz)
```
## Assigning EPSG code to a simple feature data frame

```{r echo = TRUE, eval = TRUE}
mpsz3414 <- st_set_crs(mpsz, 3414)

```

```{r echo = TRUE, eval = TRUE}
st_crs(mpsz)
```
## Transforming the projection of preschool from wgs84 to svy21

```{r echo = TRUE, eval = TRUE}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

```{r echo = TRUE, eval = TRUE}
preschool3414
```

# Importing and Converting An Aspatial Data

## Importing the Spatial data

```{r echo = TRUE, eval = TRUE}
listings <- read_csv("data/aspatial/listings.csv")
```
```{r echo = TRUE, eval = TRUE}
glimpse(listings) 
```
## Creating a simple feature data frame from an aspatial data frame

```{r echo = TRUE, eval = TRUE}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

```{r echo = TRUE, eval = TRUE}
glimpse(listings) 
```

# Geoprocessing with sf package

## Buffering

The scenario: The authority is planning to upgrade the exiting cycling path. To do so, they need to acquire 5 metres of reserved land on the both sides of the current cycling path. You are tasked to determine the extend of the land need to be acquired and their total area.

```{r echo = TRUE, eval = TRUE}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30) 
```

```{r echo = TRUE, eval = TRUE}
buffer_cycling$AREA <- st_area(buffer_cycling)
sum(buffer_cycling$AREA)
```
# Point-in-polygon count

The scenario: A pre-school service group want to find out the numbers of pre-schools in each Planning Subzone.

The code chunk below performs two operations at one go. Firstly, identify pre-schools located inside each Planning Subzone by using st_intersects(). Next, length() of Base R is used to calculate numbers of pre-schools that fall inside each planning subzone.

```{r echo = TRUE, eval = TRUE}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

```{r echo = TRUE, eval = TRUE}
summary(mpsz3414$`PreSch Count`)
```
To list the planning subzone with the most number of pre-school, the top_n() of dplyr package is used as shown in the code chunk below.

```{r echo = TRUE, eval = TRUE}
top_n(mpsz3414, 1, `PreSch Count`)
```

Calculate the density of pre-school by planning subzone.
```{r echo = TRUE, eval = TRUE}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
```

Next, mutate() of dplyr package is used to compute the density by using the code chunk below.

```{r echo = TRUE, eval = TRUE}
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

# Explorotary Data Analysis (EDA)

In this section, you will learn how to use appropriate ggplot2 functions to create functional and yet truthful statistical graphs for EDA purposes.

## Histogram
```{r echo = TRUE, eval = TRUE}
hist(mpsz3414$`PreSch Density`)
```
```{r echo = TRUE, eval = TRUE}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

## Scatter Plot

plotting a scatterplot showing the relationship between Pre-school Density and Pre-school Count

```{r echo = TRUE, fig.width=5, fig.height=5}
ggplot(data = mpsz3414, 
       aes(y = `PreSch Count`,
           x = as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
                 fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
       x = "Pre-school Density (per km sq)",
       y = "Pre-school Count")
```





