---
title: "Hands-on Exercise 3"
description: |
 In this hands-on exercise, I will learn how to plot functional and truthful choropleth maps by using an R package called tmap package.
author:
  - name: Niharika Avula 
    url: https://example.com/norajones
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Importing the required Packages

```{r echo = TRUE, eval = TRUE}
packages <- c('sf', 'tmap', 'tidyverse')
for (p in packages){
  if (!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
  }
```

# The Data

## Two data set will be used to create the choropleth map. They are:

Master Plan 2014 Subzone Boundary (Web) (i.e. MP14_SUBZONE_WEB_PL) in ESRI shapefile format. It consists of the geographical boundary of Singapore at the planning subzone level. The data is based on URA Master Plan 2014.

Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2011-2020 in csv format (i.e. respopagesextod2011to2020.csv). This is an aspatial data fie. Although it does not contain any coordinates values, but it’s PA and SZ fields can be used as unique identifiers to geocode to MP14_SUBZONE_WEB_PL shapefile.

# Importing Data

## Importing Geospatial Data 

The code chunk below uses the st_read() function of sf package to import

```{r echo = TRUE, eval = TRUE}
mpsz <- st_read(dsn = "Data/Geospatial",
                layer = "MP14_SUBZONE_WEB_PL")
```
```{r echo = TRUE, eval = TRUE}
mpsz
```

## Importing Attribute Data

```{r echo = TRUE, eval = TRUE}
popdata <- read_csv("Data/Aspatial/respopagesextod2011to2020.csv")

```

# Data Preparation

To prepare the thematic mapd, the following data table with year 2020 values is prepared. 

* YOUNG: age group 0 to 4 until age groyup 20 to 24,
* ECONOMY ACTIVE: age group 25-29 until age group 60-64,
* AGED: age group 65 and above,
* TOTAL: all age group, and
* DEPENDENCY: the ratio between young and aged against economy active group

## Data Wrangling

The following data wrangling and transformation functions will be used:

* pivot_wider() of tidyr package, and
* mutate(), filter(), group_by() and select() of dplyr package

```{r echo = TRUE, eval = TRUE}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)

```
## Combining the Attribute and Geospatial Data (Georelational Data)

This code chunk convert the values in PA and SZ fields to uppercase to match with the data in the SUBZONE_N and PLN_AREA_N

```{r echo = TRUE, eval = TRUE}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)

```

left_join() of dplyr is used to join the geographical data and attribute table using planning subzone name e.g. SUBZONE_N and SZ as the common identifier. mpsz simple feature data frame is used as the left data table to ensure that the output will be a simple features data frame.

```{r echo = TRUE, eval = TRUE}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))

```

# Choropleth Mapping Geospatial Data Using tmap

The following two approaches are explored to prepare thematic map using tmap:

* Plotting a thematic map quickly by using qtm().
* Plotting highly customisable thematic map by using tmap elements

## Plotting a choropleth map quickly by using qtm()

This method is concise and provides a good default visualisation in many cases. The code chunk below will draw a cartographic standard choropleth map as shown below.

* tmap_mode() with “plot” option is used to produce a static map. For interactive mode, “view” option should be used.
* fill argument is used to map the attribute (i.e. DEPENDENCY)

```{r echo = TRUE, eval = TRUE}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")

```
## Creating a choropleth map by using tmap’s elements

The disadvantge of qtm() is that it makes aesthetics of individual layers harder to control. To draw a high quality cartographic choropleth map as shown in the figure below, tmap’s drawing elements should be used

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))

```

The following steps show how the tmap functions that used to plot these elements

### Step 1: Drawing a Base Map

The basic building block of tmap is tm_shape() followed by one or more layer elemments such as tm_fill() and tm_polygons(). In the code chunk below, tm_shape() is used to define the input data (i.e mpsz_pop2020) and tm_polygons() is used to draw the planning subzone polygons

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```
## Step 2: Drawing a Choropleth Map using tm_polygon ()

To draw a choropleth map showing the geographical distribution of a selected variable by planning subzone, we just need to assign the target variable such as Dependency to tm_polygons().

* Default interval binning used is called "Pretty"
* Default colour scheme used is YlOrRd of ColorBrewer
* By default, missing values are shaded in grey

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```
## Step 3: Drawing a choropleth map using tm_fill() and tm_border()

tm_polygons() is a wraper of tm_fill() and tm_border(). tm_fill() shades the polygons by using the default colour scheme and tm_borders() adds the borders of the shapefile onto the choropleth map.

The code chunk below draws a choropleth map by using tm_fill() alone.

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

To add the boundary of the planning subzones, tm_borders will be used.

* alpha argument is used to define transparency number between 0 (totally transparent) and 1 (not transparent). By default, the alpha value of the col is used (normally 1)
* col = border colour
* lwd = border line width. The default is 1
* lty = border line type. The default is “solid”

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  alpha = 1)
```

# Data classification methods of tmap

The point of classification is to take a large number of observations and group them into data ranges or classes. tmap provides a total ten data classification methods, namely: fixed, sd, equal, pretty (default), quantile, kmeans, hclust, bclust, fisher, and jenks.

To define a data classification method, the style argument of tm_fill() or tm_polygons() will be used.

## Type 1: Quantile Data Classification

The code chunk below shows a quantile data classification that used 5 classes

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

## Type 2: Equal Data Classification

Note: The distribution of quantile data classification method are more evenly distributed then equal data classification method

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

## Type 3: Custome Break Data Classification

For all the built-in styles, the category breaks are computed internally. In order to override these defaults, the breakpoints can be set explicitly by means of the breaks argument to the tm_fill(). It is important to note that, in tmap the breaks include a minimum and maximum. As a result, in order to end up with n categories, n+1 elements must be specified in the breaks option (the values must be in increasing order)

Code chunk below will be used to compute and display the descriptive statistics of DEPENDENCY field in order to set the break points

```{r echo = TRUE, eval = TRUE}
summary(mpsz_pop2020$DEPENDENCY)
```

Based on the above results, breaks vector can be classified as c(0, 0.60, 0.70, 0.80, 0.90, 1.00)

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

# Colour Scheme

tmap supports colour ramps either defined by the user or a set of predefined colour ramps from the RColorBrewer package

## Using COLOURBREWER Palatte

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```
To reverse the colour shading, add a “-” prefix.

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```
# Map Layouts
In tmap, several legend options are provided to change the placement, format and appearance of the legend.

## Map Legend

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

## Map Style

tmap allows a wide variety of layout settings to be changed. They can be called by using tmap_style().

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

## Cartographic Furniture

Beside map style, tmap also also provides arguments to draw other map furniture such as compass, scale bar and grid lines. In the code chunk below, tm_compass(), tm_scale_bar() and tm_grid() are used to add compass, scale bar and grid lines onto the choropleth map

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

# Drawing Small Multiple Choropleth Maps

Small multiple maps, also referred to as facet maps, are composed of many maps arrange side-by-side, and sometimes stacked vertically. Small multiple maps enable the visualisation of how spatial relationships change with respect to another variable, such as time.

In tmap, small multiple maps can be plotted in three ways:

* by assigning multiple values to at least one of the asthetic arguments
* by defining a group-by variable in tm_facets()
* by creating multiple stand-alone maps with tmap_arrange()

## By assigning multiple values to at least one of the asthetic arguments

Small multiple choropleth maps are created by defining ncols in tm_fill()

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```
Small multiple choropleth maps are created by assigning multiple values to at least one of the aesthetic arguments

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```
## By defining a group-by variable in tm_facets()

In this example, multiple small choropleth maps are created by using tm_facets()

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

## By creating multiple stand-alone maps with tmap_arrange()

In this example, multiple small choropleth maps are created by creating multiple stand-alone maps with tmap_arrange().

```{r echo = TRUE, eval = TRUE}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```
# Mappping Spatial Object Meeting a Selection Criterion

Instead of creating small multiple choropleth map, you can also use selection funtion to map spatial objects meeting the selection criterion

```{r echo = TRUE, eval = TRUE}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```
The End