---
title: "Hands-on Exercise 5.1"
description: |
 "In this hands-on exercise, I will learn to use appropriate functions of spNetwork package: To derive network constrained kernel density estimation (NetKDE), and to perform network G-function and k-function analysis"
author:
  - name: Niharika Avula 
    url: https://example.com/norajones
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
        toc: TRUE
        toc_depth: 2
---

```{r setup, include=FALSE, eval = TRUE, echo = TRUE, message = FALSE, error = FALSE, fig.retina = 3}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Overview
Network constrained Spatial Point Patterns Analysis (NetSPAA) is a collection of spatial point patterns analysis methods special developed for analysing spatial point event occurs on or alongside network. The spatial point event can be locations of traffic accident or childcare centre for example. The network, on the other hand can be a road network or river network.

In this hands-on exercise, I will learn using appropriate functions of spNetwork package:

* to derive network constrained kernel density estimation (NetKDE), and
* to perform network G-function and k-function analysis

# 2. Data 
In this study, I will analyse the spatial distribution of childcare centre in Punggol planning area. For the purpose of this study, two geospatial data sets will be used. They are:

Punggol_St, a line features geospatial data which store the road network within Punggol Planning Area.
pg_owin, a point feature geospatial data which store the location of childcare centres within Punggol Planning Area.

# 3. Installing and launching the R packages

In this hands-on exercise, four R packages will be used, they are:

**spNetwork** which provides functions to perform Spatial Point Patterns Analysis such as kernel density estimation (KDE) and K-function on network. It also can be used to build spatial matrices (‘listw’ objects like in ‘spdep’ package) to conduct any kind of traditional spatial analysis with spatial weights based on reticular distances.

**rgdal** which provides bindings to the ‘Geospatial’ Data Abstraction Library (GDAL) (>= 1.11.4) and access to projection/transformation operations from the PROJ library. In this exercise, rgdal will be used to import geospatial data in R and store as sp objects.

**sp** which provides classes and methods for dealing with spatial data in R. In this exercise, it will be used to manage SpatialPointsDataFrame and SpatiaLinesDataFrame, and for performing projection transformation.

**tmap** which provides functions for plotting cartographic quality static point patterns maps or interactive maps by using leaflet API.

```{r}
packages = c('sp', 'rgdal', 'spNetwork', 'tmap')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}

```

# 4. Data Import and Preparation

The code chunk below uses readOGR() of rgdal package to important Punggol_St and Punggol_CC geospatial data sets as SpatialLinesDataFrame and SpatialPointsDataFrame respectively

```{r}
network <- readOGR(dsn="data/geospatial", 
                   layer="Punggol_St",
                   verbose = FALSE)
childcare <- readOGR(dsn="data/geospatial",
                     layer="Punggol_CC",
                     verbose = FALSE)
```
Code chunk below can be used to print the content of network SpatialLineDataFrame and childcare SpatialPointsDataFrame.


```{r, echo = FALSE}
str(network)
str(childcare)

```
# 5. Assign CRS to OGR data

From the above results, we can see that the Projected CRS is SVY21. Hence, we can assign the EPSG code using the following code chunk:

```{r}
childcare <-spTransform(childcare, CRS("+init=epsg:3414"))
network <- spTransform(network,CRS("+init=epsg:3414"))
```

# 6. Visualising the Geospatial Data

Before we jump into the analysis, it is a good practice to visualise the geospatial data.

There are at least two ways to visualise the geospatial data. 

## 6.1 Using plot() of Base R as shown in the code chunk below

```{r}
plot(network)
plot(childcare,add=T,col='red',pch = 19)
```
## 6.2 Using tmap package to visualise the geospatial data with high cartographic quality and interactive manner

```{r}
tmap_mode('view')
tm_shape(childcare)+
  tm_dots() +
tm_shape(network)+
  tm_lines()

```

```{r}
tmap_mode('plot')
```

# 7. Network Constrained KDE (NetKDE) Analysis

NetKDE analysis will be performed using appropriate functions provided in spNetwork package.

## 7.1 Preparing the lixels objects

* **lixelize_lines()** of spNetwork is used to cut SpatialLines object into lixels with a specified minimal distance
* Set length of a lixel, lx_length to 700m,
* Set minimum length of a lixel, mindist to 350m

```{r}
lixels <- lixelize_lines(network,700,mindist = 350)
lixels
```

Other important notes:

* After cut, if length of final lixel is shorter than minimum distance, then it is added to previous lixel
* If NULL, then mindist = maxdist/10
* Segments that are already shorter than the minimum distance are not modified

## 7.2 Generating line centre points

**lines_center()** of spNetwork will be used to generate a SpatialPointsDataFrame (i.e. samples) with line centre points. The points are located at center of the line based on the length of the line.

```{r}
samples <- lines_center(lixels)
```
# 8. Performing NetKDE

Code chunk below is used to compute the NetKDE

```{r}
densities <- nkde(network, 
                  events = childcare,
                  w = rep(1,nrow(childcare)),
                  samples = samples,
                  kernel_name = "quartic",
                  bw = 300, 
                  div= "bw", 
                  method = "simple", 
                  digits = 1, 
                  tol = 1,
                  grid_shape = c(1,1), 
                  max_depth = 8,
                  agg = 5, #we aggregate events within a 5m radius (faster calculation)
                  sparse = TRUE,
                  verbose = FALSE)
```
More on the arguments:

* **kernel_name** indicates quartic kernel is used.
  + Other possible kernel methods supported by spNetwork are: triangle, gaussian, scaled gaussian, tricube, cosine, triweight, epanechnikov or uniform.
  
* **method** indicates simple method is used to calculate the NKDE. More on the possible methods supported by spNetwork:
  + **simple:** the distances between events and sampling points are replaced by network distances, and the formula of the kernel is adapted to calculate the density over a linear unit instead of an areal unit.
  + **discontinuous:** Equally “divides” the mass density of an event at intersections of lixels.
  + **continuous:** Divide the mass of the density at intersection but adjusts the density before the intersection to make the function continuous
  
# 9. Visualising NetKDE

## 9.1 Insert the computed density values (i.e. densities) into samples and lixels objects

```{r}
samples$density <- densities
lixels$density <- densities

```


## 9.2 Rescale density values

Since svy21 projection system is in meter, the computed density values are very small i.e. 0.0000005. The code chunk below is used to resale the density values from number of events per meter to number of events per kilometer.

```{r}
samples$density <- samples$density*1000
lixels$density <- lixels$density*1000
```

## 9.3 Plot map

The code below uses appropriate functions of tmap package to prepare interactive and high cartographic quality map visualisation.

```{r}
tmap_mode('view')
tm_shape(lixels)+
  tm_lines(col="density")+
tm_shape(childcare)+
  tm_dots()

```
```{r}
tmap_mode('plot')
```
We see from map above:

* Road segments with relatively higher density of childcare centres (darker color)
* Road segments with relatively lower density of childcare centres (lighter color)

# 10. Network Constrained G- and K-Function Analysis

* Here, we will perform complete spatial randomness (CSR) test by using *kfunctions()* of **spNetwork package.**

* The null hypothesis is defined as: 
  + **Ho:** The observed spatial point events (i.e distribution of childcare centres) are UNIFORMLY DISTRIBUTED over a street network in Punggol Planning Area.

* The CSR test is based on the assumption of the binomial point process which implies the hypothesis that the childcare centres are randomly and independently distributed over the street network.

* If this hypothesis is rejected, we may infer that the distribution of childcare centres are spatially interacting and dependent on each other; as a result, they may form nonrandom patterns.

```{r}
kfun_childcare <- kfunctions(network, 
                             childcare,
                             start = 0, 
                             end = 1000, 
                             step = 50, 
                             width = 50, 
                             nsim = 50, 
                             resolution = 50,
                             verbose = FALSE, 
                             conf_int = 0.05)

```

More on the arguments used above:

* lines: A SpatialLinesDataFrame with the sampling points. The geometries must be a SpatialLinesDataFrame (may crash if some geometries are invalid).
* points: A SpatialPointsDataFrame representing the points on the network. These points will be snapped on the network.
* start: A double, the start value for evaluating the k and g functions.
* end: A double, the last value for evaluating the k and g functions.
* step: A double, the jump between two evaluations of the k and g function.
* width: The width of each donut for the g-function.
* nsim: An integer indicating the number of Monte Carlo simulations required. In the above example, 50 simulation was performed. Note: most of the time, more simulations are required for inference
* resolution: When simulating random points on the network, selecting a resolution will reduce greatly the calculation time. When resolution is null the random points can occur everywhere on the graph. If a value is specified, the edges are split according to this value and the random points are selected vertices on the new network.
* conf_int: A double indicating the width confidence interval (default = 0.05)

The output of kfunctions() is a list with the following values:

* plotkA, a ggplot2 object representing the values of the k-function
* plotgA, a ggplot2 object representing the values of the g-function
* valuesA, a DataFrame with the values used to build the plots

## 10.1 Visualise the ggplot2 object of k-function

```{r}
kfun_childcare$plotk
```
From the plot above:

* **blue line** is the empirical network K-function of the childcare centres in Punggol planning area

* **gray envelop** represents the results of the 50 simulations in the interval 2.5% - 97.5%

* Because **blue line** between the distance of 250m-400m are **below the gray area**, we can infer that the **childcare centres in Punggol planning area resemble regular pattern at the distance of 250m-400m.**

## 10.2 Visualise the values object of k-function

```{r}
kfun_childcare$values
```