---
title: "In-Class Exercise 4"
description: |
  In this in-class exercise, I will learn how to apply spatial point analysis techniques
  
author:
  - name: Niharika Avula 
    url: https://example.com/norajones
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# R Packages

The following five R Packages are used
* sf is mainly used for importing geospatial data
* Maptools is used to convert spatial object to ppp format (vector data conversion method)
* All output of spatstat is going to be raster grid
* sf is specially design for sf layr, raster is specially designed for raster layer

```{r}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap', 'readr')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
```

## Importing Geospatial Data

```{r echo = TRUE, eval = TRUE}
sg_sf <- st_read(dsn = "data/shapefile", layer = "CostalOutline")
mpsz_sf <- st_read(dsn = "data/shapefile", layer = "MP14_SUBZONE_WEB_PL")
```
## Importing Aspatial Data

```{r echo = TRUE, eval = TRUE}
childcare <- read_rds("data/rds/childcare.rds")
CHAS <- read_rds("data/rds/CHAS.rds")
```

## Converting from apatial to geospatial
```{r echo = TRUE, eval = TRUE}
CHAS_sf <- st_as_sf(CHAS, 
                    coords = c("X_COORDINATE", 
                               "Y_COORDINATE"), 
                    crs = 3414)
```

```{r echo = TRUE, eval = TRUE}
childcare$Lat <- as.numeric(childcare$Lat)
childcare$Lng <- as.numeric(childcare$Lng)
childcare_sf <- st_as_sf(childcare, 
                    coords = c("Lng", 
                               "Lat"), 
                       crs = 4326) %>%
  st_transform(crs = 3414)
```

## Geospatial Data Wrangling

### Converting from sf to Spatial* Classes

```{r echo = TRUE, eval = TRUE}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
sg <- as_Spatial(sg_sf)
```

### Converting from Spatial* Classes to sp format

```{r echo = TRUE, eval = TRUE}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
SG_sp <- as(sg, "SpatialPolygons")
```

### Converting from sp format to ppp format
```{r echo = TRUE, eval = TRUE}
childcare_ppp <- as(childcare_sp, "ppp")
CHAS_ppp <- as(CHAS_sp, "ppp")
```

```{r echo = TRUE, eval = TRUE}
tmap_mode ('view')
tm_shape(childcare_sf)+
  tm_dots(alpha = 0.4,
          col = "blue",
          size = 0.05) +
  tm_shape(CHAS_sf) +
  tm_dots(alpha = 0.4,
          col = "red",
          size = 0.05)

```

```{r echo = TRUE, eval = TRUE}
tmap_mode('plot')
```