---
title: "In-Class Exercise 6"
description: |
  In this exercise, I will learn how to derive spatial weights by using functions provided by spdep package. I will also learn how to apply these spatial weights to compute spatially lagged values
author:
  - name: Niharika Avula
    url: https://example.com/norajones
date: "`r Sys.Date()`"
output: distill::distill_article
   self_contained: false
---

```{r setup, include=FALSE,eval = TRUE, echo = TRUE, message = FALSE, error = FALSE, fig.retina = 3}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

In this exercise, I will learn how to compute spatial weights using R:

* Import geospatial data using appropriate function(s) of **sf** package,
* Import csv file using appropriate function of **readr** package,
* Perform relational join using appropriate join function of **dplyr** package,
* Compute spatial weights using appropriate functions of **spdep** package, and
* Calculate spatially lagged variables using appropriate functions of **spdep** package

# Data

Two data sets will be used in this exercise, they are:

* Hunan county boundary layer. This is a geospatial data set in ESRI shapefile format.
* Hunan_2012.csv: This csv file contains selected Hunan’s local development indicators in 2012.

# Getting Started

The code chunk below installs the following packages 

```{r}
packages = c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```
# Getting the Data Into R Environment

The geospatial data is in ESRI shapefile format and the attribute table is in csv fomat, so they will be imported into the R environment in this section.

## Import shapefile into R environment
The code chunk below uses st_read() of sf package to import Hunan shapefile into R. The imported shapefile will be simple features Object of sf.

```{r}
hunan <- st_read(dsn = "data/shapefile", 
                 layer = "Hunan")

```

## Import csv file into R environment

The code chunk below uses read_csv() of readr package to import Hunan_2012.csv into R.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

