---
title: "In-class Exercise 3"
description: "In this hands-on exercise, I will learn how to plot Percentile, Box, Rate maps by using an R package called tmap package."
author:
  - name: Niharika Avula
    url: https://example.com/norajones
date: 08-23-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r echo = TRUE, eval = TRUE}
packages <- c('tmap', 'sf', 'tidyverse')
for (p in packages){
  if (!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
  }
```


# The Data

## Two data set will be used to create the choropleth map. They are:

Master Plan 2014 Subzone Boundary (Web) (i.e. MP14_SUBZONE_WEB_PL) in ESRI shapefile format. It consists of the geographical boundary of Singapore at the planning subzone level. The data is based on URA Master Plan 2014.

Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2011-2020 in csv format (i.e. respopagesextod2011to2020.csv). This is an aspatial data fie. Although it does not contain any coordinates values, but it’s PA and SZ fields can be used as unique identifiers to geocode to MP14_SUBZONE_WEB_PL shapefile.

# Importing Data

## Importing Geospatial Data 

The code chunk below uses the st_read() function of sf package to import

```{r echo = TRUE, eval = TRUE}
mpsz <- st_read(dsn = "Data/Geospatial",
                layer = "MP14_SUBZONE_WEB_PL")
```
```{r echo = TRUE, eval = TRUE}
mpsz
```

## Importing Attribute Data

```{r echo = TRUE, eval = TRUE}
popdata <- read_csv("Data/Aspatial/respopagesextod2011to2020.csv")

```

# Data Preparation

To prepare the thematic mapd, the following data table with year 2020 values is prepared. 

* YOUNG: age group 0 to 4 until age groyup 20 to 24,
* ECONOMY ACTIVE: age group 25-29 until age group 60-64,
* AGED: age group 65 and above,
* TOTAL: all age group, and
* DEPENDENCY: the ratio between young and aged against economy active group

## Data Wrangling

The following data wrangling and transformation functions will be used:

* pivot_wider() of tidyr package, and
* mutate(), filter(), group_by() and select() of dplyr package

```{r echo = TRUE, eval = TRUE}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)

```
## Combining the Attribute and Geospatial Data (Georelational Data)

This code chunk convert the values in PA and SZ fields to uppercase to match with the data in the SUBZONE_N and PLN_AREA_N

```{r echo = TRUE, eval = TRUE}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)

```

left_join() of dplyr is used to join the geographical data and attribute table using planning subzone name e.g. SUBZONE_N and SZ as the common identifier. mpsz simple feature data frame is used as the left data table to ensure that the output will be a simple features data frame.

```{r echo = TRUE, eval = TRUE}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))

```

# Data Visualisation

## Boxplot - Visualising Extreme Values

Boxplot is one one the popular Exploratory Data Analysis (EDA) technique used to show the statistics and distribution of data values

```{r echo = TRUE, eval = TRUE}
ggplot(data=mpsz_pop2020, 
       aes(x = "", 
           y = AGED)) +
  geom_boxplot()
```

Despite its usefulness, boxplot is not able to reveal the spatial distribution of these outliers.

## Percentile Map

Special type of quantile map with six specific categories: 0-1%,1-10%, 10-50%,50-90%,90-99%, and 99-100%.

* The corresponding breakpoints can be derived by means of the base R quantile command, passing an explicit vector of cumulative probabilities as c(0,.01,.1,.5,.9,.99,1)

This code chunk excludes records with NA
```{r echo = TRUE, eval = TRUE}
mpsz_pop2020a <- mpsz_pop2020 %>%
  drop_na()
```

```{r echo = TRUE, eval = TRUE}
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- mpsz_pop2020a["DEPENDENCY"] %>%
  st_set_geometry(NULL)
quantile(var[,1], percent)
```
The code chunk below is an R function to extract a variable as a vector out of an sf data frame

* Arguments: vname: variable name (as character, in quotes); df: name of sf data frame
* returns: v: vector with values (without a column name)

```{r echo = TRUE, eval = TRUE}
get.var <- function(vname,df) {
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

```{r echo = TRUE, eval = TRUE}
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- get.var("DEPENDENCY", mpsz_pop2020a)
bperc <- quantile(var,percent)
tm_shape(mpsz_pop2020) +
  tm_polygons() +
tm_shape(mpsz_pop2020a) +
  tm_fill("DEPENDENCY",
          title="DEPENDENCY",
          breaks=bperc,
          palette="Blues",
          labels=c("< 1%", "1% - 10%",
                   "10% - 50%", 
                   "50% - 90%",
                   "90% - 99%", 
                   "> 99%"))  +
  tm_borders() +
  tm_layout(title = "Percentile Map", 
            title.position = c("right",
                               "bottom"))
```

## Percentile Mapping Function
```{r echo = TRUE, eval = TRUE}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam,df)
  bperc <- quantile(var,percent)
  tm_shape(mpsz_pop2020) +
  tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(title = mtitle, title.position = c("right","bottom"))
}
```

```{r echo = TRUE, eval = TRUE}
youngmap <- percentmap("YOUNG", mpsz_pop2020a)
agedmap <- percentmap("AGED", mpsz_pop2020a)
```

## Box Map

* Displaying summary statistics on a choropleth map by using the basic principles of boxplot
* To create a box map, a custom breaks specification will be used. However, the break points for the box map vary depending on whether lower or upper outliers are present.

The code chunk below is an R function that creates break points for a box map.

Arguments:
* v: vector with observations
* mult: multiplier for IQR (default 1.5)
Returns:
* bb: vector with 7 break points compute quartile and fences

```{r echo = TRUE, eval = TRUE}
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```

```{r echo = TRUE, eval = TRUE}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

```{r echo = TRUE, eval = TRUE}
var <- get.var("AGED", mpsz_pop2020a) 
boxbreaks(var)
```

The code chunk below is an R function to create a box map.

Arguments:
* vnam: variable name (as character, in quotes)
* df: simple features polygon layer
* legtitle: legend title
* mtitle: map title
* mult: multiplier for IQR

Returns:
* a tmap-element (plots a map)

```{r echo = TRUE, eval = TRUE}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(title = mtitle, 
            title.position = c("right",
                               "bottom"))
}
```
Plotting the box map

```{r echo = TRUE, eval = TRUE}
boxmap("ECONOMY ACTIVE", mpsz_pop2020a)
```

The box map reveals that there are six upper outliers (i.e. planning subzone with extremely high numbers of aged population). Four of the upper outliers are located at the eastern region and they are closed to each others and there is no lower outlier.