---
title: "In-Class Exercis 5"
description: |
 "In this hands-on exercise, I will learn spatial point analysis techniques"
author:
  - name: Nora Jones 
    url: https://example.com/norajones
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE, eval = TRUE, echo = TRUE, message = FALSE, error = FALSE, fig.retina = 3}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing and Loading the R package

* maptools is for creating spatial objects
* sf is for importing geospatial data
* rater is for creating raster objects
* spatstat is for performing spatial point pattern analysis
* tmap is for plotting thematic maps
* tidyverse is for data handling
* plotly is used to plot interactive plots
* ggthemes is an extension of ggplot

```{r}
packages <- c('maptools', 'sf', 'raster', 'spatstat', 'tmap', 'tidyverse', 'plotly', 'ggthemes')

for (p in packages){
  if (!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
  }

```
## Importing the Geospatial Data

Impoting shapefile using *st_read()* of *sf* package. The output object is in tibble sf object class

``` {r}
mpsz_sf <- st_read(dsn = "data/shapefile", 
                   layer = "MP14_SUBZONE_WEB_PL")
```

Projection is in svy21

## Importing Geospatial Data 

``` {r}
mpsz_sf <- st_read(dsn = "data/shapefile", 
                   layer = "MP14_SUBZONE_WEB_PL")
```

## Importing Aspatial Data from mrds folder

read_rds () of readr package is used instead of readRDS() of base R is used. This is because output of read_rds() is in tibble object.


``` {r}

childcare <- read_rds("data/rds/childcare.rds") 
CHAS <- read_rds("data/rds/CHAS.rds") 

```

Note that there are some data issue in childcare data frame because `Lat` and `Lng` should be in numeric data type. The coordinate fields seem to be in decimal degrees, Hence, `wgs84` referencing system is assumed.

## Converting the Aspatial data frame into sf objects


```{r}
CHAS_sf <- st_as_sf(CHAS, 
                    coords = c("X_COORDINATE",
                               "Y_COORDINATE"),
                    crs = 3414)
```

Note: st_as_sf accept coordinates in character data type

```{r}
childcare_sf <- st_as_sf(childcare, 
                    coords = c("Lng",
                               "Lat"),
                    crs = 4326) %>%
  st_transform(crs = 3414)
```

## Plotting for reviewing

```{r}
tmap_mode("view")
tm_shape(childcare_sf) +
  tm_dots(alpha = 0.4, 
          col = "blue",
          size = 0.05) +
  tm_shape(CHAS_sf) +
  tm_dots(alpha = 0.4, 
          col = "red",
          size = 0.05)

```

## Geospatial Data Wrangling 

### Converting from sf to spatial* data frame classes

as_Spatial () of sf package

```{r}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <-  as_Spatial (mpsz_sf)
```

### Converting from spatial* data frame into Spatial* Objects

as.SpatialPoints() or as.SpatialPolygons() of maptools package

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
mpsz_sp <- as(mpsz, "SpatialPolygons")
```

### Converting from Spatial* Objects into ppp objects

Using as.ppp() of maptools package; lost all the projection and reference data, Whatever that is going to remain will only have x and y coordinates

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
CHAS_ppp <- as(CHAS_sp, "ppp")
```

### Removing duplicate points using jitter function

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry = TRUE,
                             nsim = 1,
                             drop = TRUE)

any(duplicated(childcare_ppp_jit))
```
```{r}
CHAS_ppp_jit <- rjitter(CHAS_ppp, 
                             retry = TRUE,
                             nsim = 1,
                             drop = TRUE)

any(duplicated(CHAS_ppp_jit))
```
Since the data type is changed, we cannot use tmap to plot/view the data

## Extracting Punggol Planning Area

There is a comma behind 

```{r}
pg <- mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
```

### Converting SpatialPolygonDataFrame into SpatialPolygon object

```{r}
pg_sp <- as(pg, "SpatialPolygons")

```

### Converting spatialPolygons into owin object

owin is a unique object that 

```{r}
pg_owin <- as(pg_sp, "owin")
```

### Extract spatial points within OWIN

```{r}
childcare_pg <- childcare_ppp_jit[pg_owin]
CHAS_pg <- CHAS_ppp_jit[pg_owin]
```

```{r}
plot(childcare_pg)
```

### L-Function

```{r}

L_childcare <- envelope(childcare_pg,
                        Lest,
                        nsim = 99,
                        rank = 1,
                        global = TRUE)
```



```{r}
title <- "Pairwise Distance: L function"

Lcsr_df <- as.data.frame(L_childcare)

colour=c("#0D657D","#ee770d","#D3D3D3")
csr_plot <- ggplot(Lcsr_df, aes(r, obs-r))+
  # plot observed value
  geom_line(colour=c("#4d4d4d"))+
  geom_line(aes(r,theo-r), colour="red", linetype = "dashed")+
  # plot simulation envelopes
  geom_ribbon(aes(ymin=lo-r,ymax=hi-r),alpha=0.1, colour=c("#91bfdb")) +
  xlab("Distance r (m)") +
  ylab("L(r)-r") +
  geom_rug(data=Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,], sides="b", colour=colour[1])  +
  geom_rug(data=Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,], sides="b", colour=colour[2]) +
  geom_rug(data=Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,], sides="b", color=colour[3]) +
  theme_tufte()+
  ggtitle(title)

text1<-"Significant clustering"
text2<-"Significant segregation"
text3<-"Not significant clustering/segregation"

# the below conditional statement is required to ensure that the labels (text1/2/3) are assigned to the correct traces
if (nrow(Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,])==0){ 
  if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text3, traces = 4) %>%
      rangeslider() 
  }else if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      rangeslider() 
  }else {
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider() 
  }
} else if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){
  if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      rangeslider() 
  } else{
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider()
  }
} else{
  ggplotly(csr_plot, dynamicTicks=T) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    rangeslider()
  }
```

```{r}

L_CHAS <- envelope(CHAS_pg,
                        Lest,
                        nsim = 99,
                        rank = 1,
                        global = TRUE)
```


```{r}
title <- "Pairwise Distance: L function"

Lcsr_df <- as.data.frame(L_CHAS)

colour=c("#0D657D","#ee770d","#D3D3D3")
csr_plot <- ggplot(Lcsr_df, aes(r, obs-r))+
  # plot observed value
  geom_line(colour=c("#4d4d4d"))+
  geom_line(aes(r,theo-r), colour="red", linetype = "dashed")+
  # plot simulation envelopes
  geom_ribbon(aes(ymin=lo-r,ymax=hi-r),alpha=0.1, colour=c("#91bfdb")) +
  xlab("Distance r (m)") +
  ylab("L(r)-r") +
  geom_rug(data=Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,], sides="b", colour=colour[1])  +
  geom_rug(data=Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,], sides="b", colour=colour[2]) +
  geom_rug(data=Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,], sides="b", color=colour[3]) +
  theme_tufte()+
  ggtitle(title)

text1<-"Significant clustering"
text2<-"Significant segregation"
text3<-"Not significant clustering/segregation"

# the below conditional statement is required to ensure that the labels (text1/2/3) are assigned to the correct traces
if (nrow(Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,])==0){ 
  if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text3, traces = 4) %>%
      rangeslider() 
  }else if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      rangeslider() 
  }else {
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider() 
  }
} else if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){
  if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      rangeslider() 
  } else{
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider()
  }
} else{
  ggplotly(csr_plot, dynamicTicks=T) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    rangeslider()
  }
```